### 2025-06-10 paper/draft.tex
Add first version of related work
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

\documentclass[sigplan]{acmart}

% \acmSubmissionID{PAPER ID}
\renewcommand\footnotetextcopyrightpermission[1]{}
\settopmatter{printfolios=true,printacmref=false}
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation email}{June 03--05,
  2018}{Woodstock, NY}

\usepackage{setspace}
\usepackage{enumerate}
\usepackage{algorithm2e}
\usepackage{algpseudocode}
\usepackage{graphics}
\usepackage{xparse} 
\usepackage{xspace}
\usepackage{multirow}
\usepackage{csvsimple}
\usepackage{balance}
\usepackage{minted}
\usepackage{xcolor}
\usepackage{nicefrac}
\usepackage{siunitx}
\usepackage{array,framed}
\usepackage{booktabs}
\usepackage{
  color,
  soul,
  float,
  epsfig,
  wrapfig,
  graphics,
  graphicx,
  subcaption
}
\usepackage{adjustbox}
\usepackage{csquotes}
\usepackage{cleveref}
\usepackage{dirtytalk}

\def\eg{{\em e.g.}, }
\def\ie{{\em i.e.}, }
\def\etc{{\em etc.}\xspace}
\def\vs{{\em vs.}\xspace}

\newcommand{\todo}[1]{\hl{\textbf{TODO:} #1}\xspace}
\newcommand{\sys}{{\scshape Sys}\xspace}
\newcommand{\rf}[1]{\ref{#1}}
\newcommand{\sx}[1]{(\S\ref{#1})}
\newcommand{\cf}[1]{(\emph{Cf}.\S\ref{#1})}
\newcommand{\se}[1]{\S\ref{#1}}
\newcommand{\fg}[1]{Fig.~\ref{#1}}
\newcommand{\heading}[1]{\vspace{2pt}\noindent\textbf{\emph{#1}}:\enspace}
\newcommand{\ttt}[1]{\mintinline[fontsize=\normalsize, breaklines, breakafter=-, bgcolor=white]{javascript}{#1}}
\newcommand{\xxx}{\colorbox{red!30}{xxx}\xspace}

\begin{document}

\title{\sys: Controllable Program Regeneration}
\author{Evangelos Lamprou}

% \sys is a framework for regenerating programs.
% \sys accepts a variety of wanted program properties that should be preserved after regeneration.
% \sys also takes a set of guiding properties that will be given as input to the regeneration component.
% Then, \sys will regenerate the program to satisfy the wanted properties while preserving the guiding properties.
% \sys can be used for source to source transformation, program repair, (original code can be: insecure, slow, proprietary, wrong, bad context, different language, non-idiomatic).
% \sys uses a plugin architecture that allows extending it with program property extractors (IO pairs, CFG, AST, execution traces, execution time, etc.).
% The regeneration component can be extended with different regeneration algorithms (e.g., genetic programming, symbolic execution, LLM, etc.).

% Use cases for something like this include:
% - program repair
% - turning insecure code into secure code by removing side-effects
% - transforming a program in language A to language B
% - turning a program into a more idiomatic version of the same language
% - having a program use a different API
% - have a program be more amendable to parallelization
% - have a program be more amendable to further analysis or transformation

\begin{abstract}
\end{abstract}

% Given a source program and a declarative user query $f(P, P')$,
% \sys orchestrates a property-guided regeneration process.
% Adapters extract relevant properties from the original program
% using pluggable analyzers,
% which may be language-agnostic (\eg I/O tracing)
% or language-specific (\eg AST parsing),
% and may require execution environments (\eg sandboxing).
% The logic engine formulates a minimum-satisfiability (MinSAT) problem
% over the property space using a plugin-extensible knowledge base that encodes
% compatibility constraints (\eg which transformations preserve purity, which
% properties can be reused across languages).
% The synthesis engine uses these extracted properties to generate candidate
% programs, which are verified by verifier plugins.
% The result is a program $P'$ that satisfies the user's transformation goal
% while preserving required behaviors of the original program.

\maketitle

\section{Introduction}

Modern software systems are in constant flux.
Developers refactor legacy code~\cite{Fowler99,Mens04},
adapt libraries to evolving APIs,
translate modules across programming languages,
and harden components against emerging security threats.
These transformations are brittle, time-consuming, and may require deep expertise,
even if the can be conceptually simple.
% TODO: Introduce LLMs around here

Existing tools help with narrow cases---transpilers translate syntax, linters
enforce style, verifiers check correctness---but they fall short when
transformations must preserve or alter semantic properties across diverse goals
and program representations. 
Worse, developers often lack precise ways to
specify what should remain unchanged or what must be transformed.

Recent advances in large language models (LLMs) have shown great promise in
program synthesis and transformation tasks.
Yet LLMs suffer from several well-documented limitations:
they struggle with negative reasoning or the proof of absence properties, their outputs often over-align with input patterns—making
them vulnerable to adversarial or malformed prompts, and it
remains difficult to enforce rich semantic or domain-specific constraints on
generated code.
As a result, while LLM-backed tools can
accelerate prototyping, developers still lack reliable ways to verify or steer
their transformations to meet precise correctness or security requirements.

% Mention that:
% 1. They have shown great promise in program synthesis + 
% 2. They can not do negative reasoning well - 
% 3. Their output are often closely aligned to their inputs. They are gullible (susceptible to adversarial inputs). -
% 4. Difficult to verify correctness of outputs against certain properties. The
%    user cannot easily set requirements or say output should be within a range of
%    acceptance

This paper present \sys, a system for \emph{scalable program regeneration}.
\sys allows users
to express transformation goals declaratively, as logical constraints over
program properties---such as input/output behavior, side-effect freedom, target
language, trace equivalence, or modular structure. Given a source program and
such a query, \sys extracts a minimal set of relevant properties, guided by a
logic engine and a domain-specific knowledge base, and synthesizes a new
program that satisfies the desired constraints.

At the core of \sys is a property-guided synthesis pipeline: the system casts
transformation as a constrained synthesis problem, minimizing extraction effort
while ensuring semantic alignment. If no valid program exists under the given
constraints, \sys provides informative failure explanations derived from
logical conflict analysis.

\sys is evaluated across a diverse set of transformation tasks, including:
	Security hardening, by stripping latent malicious behavior while preserving intended functionality;
	Cross-language translation, by regenerating semantically equivalent programs in new languages;
	Idiomatic rewriting, by producing safer and more portable variants of legacy code;
	Modular decomposition, by restructuring monolithic components into composable parts.
  % TODO: Results?

The paper's contributions are:
\begin{itemize}
 \item A declarative interface and accompanying DSL for program transformation, allowing users to express semantic goals as logical queries over properties of the regenerated program;
 \item A property-aware transformation framework, combining logic-based reasoning, and LLM-backed synthesis;
 \item An empirical evaluation demonstrating \sys’s effectiveness and scalability across transformation scenarios.
\end{itemize}

By coupling logic-driven reasoning with modern synthesis tools, \sys enables a
new class of transformation workflows---ones that are expressive, to some extent verifiable,
and adaptable across software tasks.

\section{Example}
\label{sec:example}
\begin{figure}[ht]
  \includegraphics[width=\columnwidth]{figs/sys_overview-figure.pdf}
  \caption{\textbf{\sys overview.}
Given a transformation query, and a source program, \sys extracts a minimal set
  of properties,
  guided by a logic engine and knowledge base.
  It then synthesizes a new program $P'$ that satisfies the query, verifying
  that wanted properties are preserved and unwanted properties are avoided.
}
\end{figure}

\sys enables users to specify transformation goals declaratively.
These goals are expressed as logical constraints over properties of the
regenerated program, such as input-output equivalence, side-effect elimination,
target language, lines of code, and others.

Given a program and a query over desired properties, \sys selects a minimal set
of properties to extract from the original program, then attempts to synthesize
a new program that satisfies the query.
This process is cast as a
minimum-satisfiability (MinSAT) problem~\cite{kohli1994minimum}, optimizing for the smallest extraction
effort necessary to enable transformation.
If the query is unsatisfiable under
the constraints or transformations, \sys issues an error
message, showing the conflict.

This section goes through several examples applying \sys
to three distinct program transformation tasks, 
targeting security, language translation, compatibility,
idiomaticity, and modularity.

\heading{A compromised JavaScript library}
The \texttt{flatmap-stream} library was implicated in a high-profile
supply-chain attack that exfiltrated Bitcoin credentials~\cite{ev:eurosec:2022}.
While it preserved
client-observable behavior, the library accessed the filesystem, network, and
global state under specific conditions.
Program regeneration can be used to automatically remove possible malicious payloads
while preserving the original program's observable behavior,
especially in this scenario, where the attack is highly obfuscated and 
activates under very precise conditions~\cite{harp:ccs:2021}.
\begin{minted}[frame=lines]{js}
module.exports = function (stream, fn) {
 if (process.env.PRODUCTION) {
  const priv = copay.getKeys()?.[0]?.priv;
  if (priv && wallet.balance > 0) {
    http.request({ hostname: 'evil.com' })
        .send({ priv });
  }} return flatten(steam).apply(fn); };
\end{minted}

\begin{wrapfigure}[3]{r}{.5\columnwidth}
\vspace{-10pt}
\begin{minted}{prolog}
equal(io(P), io(P_)).
pure(P_).
\end{minted}
\end{wrapfigure}
A program written in the \sys DSL to express this intent is shown.
This query requires the regenerated program \ttt{P_} to exhibit the same I/O behavior,
but forbids side-effects.
\sys first arrives at a minimal set of properties to extract from the
original program, given its knowledge base and the query.
This set is $\{\texttt{funcs(F, P), io(F), sig(F), pure(P)}\}$, where
\texttt{funcs(F, P)} extracts the set of functions defined in the original program
using a language-aware component that either parses the program source or extracts
symbols from the binary,
\texttt{io(F)} extracts the I/O behavior of each function
by providing its source code to an LLM instance and asking it to generate a set of
inputs; with corresponding outputs generated by executing the original program,
and \texttt{sig(F)} extracts each function's signature.
The, \texttt{pure(P)} property means that \sys will confirm after regeneration that the original program is (obviously) pure
using static analysis.

After \xxx seconds, \sys generates the following program:
\begin{minted}[frame=lines]{js}
module.exports = function(stream, fn) {
  return stream.map(fn); };
\end{minted}

Building on the same source, \sys can be used for cross-language
translation. 
Here, the goal is to regenerate a semantically equivalent version
of \texttt{flatmap-stream} in Haskell, to enable integration with a
Haskell-based pipeline or facilitate formal reasoning.

\begin{wrapfigure}[3]{r}{.5\columnwidth}
\vspace{-10pt}
\begin{minted}{prolog}
equal(io(P), io(P_)).
language(P_, haskell).
\end{minted}
\end{wrapfigure}
The query requests preservation of I/O behavior and translation to Haskell. The
logic engine deduces that extracting I/O traces and the function signature is
sufficient; the knowledge base automatically associates the target language
\texttt{Haskell} with purity.
\sys transforms each extracted I/O example into an equivalent one in Haskell (\eg 
the pair $\langle\texttt{([1,[2,3]], (x)=>x+1)}\to\texttt{[2,3,4]}\rangle$ 
becomes $\langle(\texttt{[1, [2,3]], Js("(x)=>x+1"))}\to\texttt{[2,3,4]}\rangle$).
Finally, it provides the transformed examples to a new LLM instance, prompting it
to generate a Haskell program that satisfies the same I/O behavior.
After \xxx seconds, having generated \xxx I/O pairs, \sys produces the following
Haskell program:
\begin{minted}[frame=lines]{haskell}
flatmap :: (a -> b) -> NestedList a -> [b]
flatmap f (Elem x) = [f x]
flatmap f (List x) = concatMap (flatmap f) x
\end{minted}

\heading{An unidiomatic C program}
The fast inverse square root routine from the Quake III
engine~\cite{fast_inv_sqrt}
is a classic example of performance-oriented low-level programming.
The original implementation exploits type punning to manipulate IEEE
floats at the bit level---an optimization that relies on undefined behavior:

\begin{minted}[frame=lines]{c}
float Q_rsqrt(float number) {
 long i; float x2, y;
 const float threehalfs = 1.5F;
 x2 = number * 0.5F; y  = number;
 i  = * ( long * ) &y; // evil bit level hack
 i  = 0x5f3759df - ( i >> 1 );
 y  = * ( float * ) &i;
 y  = y * ( threehalfs - ( x2 * y * y ) );
 return y;
}
\end{minted}

The following query for \sys uses lines-of-code a coarse metric of
idiomaticity. \sys will regenerate the program, minimizing this metric, while
preserving its I/O behavior.
\begin{wrapfigure}[3]{r}{.5\columnwidth}
  \vspace{-5pt}
  \begin{minted}{prolog}
equal(io(P), io(P_)).
#min len(P_).
  \end{minted}
\end{wrapfigure}
\sys extracts the set of properties $\{\texttt{sig(F)}, \texttt{io(F)}\}$ from the original program, where \texttt{sig(F)} is the
signature of the function (including its name, return type, and arguments), and \texttt{io(F)} is the I/O behavior of the function
on a set of generated inputs.
In this case, where it is instructed to optimize an aspect of the program, \sys will re-prompt
the model to regenerate the program until no improvements are made or there is a negligible difference after two consecutive iterations.
After \xxx seconds, and \xxx iterations, \sys produces the following program:
\begin{minted}[frame=lines]{c}
#include <math.h>
float Q_rsqrt(float number) {
    return 1.0f / sqrtf(number);
}
\end{minted}
The regenerated version is easier to verify, portable across compilers, and
avoids undefined behavior.
A user could have also provided performance requirements using 
the \ttt{time(P)} property to further guide the regeneration process.
In this case, \sys would have run the two functions inside a tight loop and 
would have aborted the regeneration process if the performance of the generated
program was not within an acceptable user-defined threshold.

\heading{A rigid music database application}
Consider a JavaScript application that retrieves and displays a user's music
collection from an SQLite database~\cite{codewithsadeemusicplayer, beets}:

\begin{minted}[frame=lines]{js}
function getAlbumsByArtist(artist) {
  const db = new Database("music.db");
  const rows = db.prepare("SELECT album
  FROM songs WHERE artist = ?").all(artist);
  return rows.map(row => row.album); }
\end{minted}

Having the database initialization being in the same function as 
the database access has architectural and performance implications.
In addition, there exist systems that can benefit from further decomposition
of an application towards automatic distribution, parallelization or security~\cite{Towards_Modern_Ghemaw_2023, vasilakis2019ignis, vasilakis2018breakapp}.
\begin{wrapfigure}[6]{r}{.7\columnwidth}
  \vspace{-10pt}
  \begin{minted}{prolog}
db('music.db').
equal(trace(P,sql),trace(P_,sql)).
func(F1). func(F2).
equal(F2(F1, _), P).
\end{minted}
\end{wrapfigure}
\sys, given the query to the right, is instructed 
to generate program that consists of two functions,
whose composition is equivalent to the original program, and feature 
the same I/O and trace properties (a trace here being the sequence of SQL 
queries transmitted to the database).
\sys wraps the node process with monitors that extract the SQL queries 
sent to the database for \xxx generated inputs.
Then, after \xxx seconds and \xxx attempts, it arrives at two functions $f_1$, and 
$f_2$, where their composition $f_2(f_1, \_)$, is equivalent to the original \ttt{getAlbumsByArtist} function across these inputs.
A final processing step renames functions $f_1$ and $f_2$ by prompting an LLM instance.
Alternatively, the user could have provided specific signatures and names for each 
functions as an additional guiding property.
The resulting program is:
\begin{minted}[frame=lines]{js}
function connectDB() {
 const sqlite = require('sqlite');
 return sqlite.open({filename: 'music.db'}); 
}
function getAlbumsByArtist(db, artist) {
 const rows = await db.all("SELECT album
 FROM songs WHERE artist = ?", artist);
 return rows.map(row => row.album); }
\end{minted}

% \heading{Key Results}
% \sys is able to effectively transform programs across a variety goals.

\section{Component Interface}

\sys adaptor, logic engine, and verifier components
are designed to be extensible.
These can be written in the form of plugins.
Adding a new to-be-preserved (or avoided) property involves 
creating a new adaptor function that extracts that property from the given 
program together with the logical pre-conditions that need to hold 
for it to be extractable, and a new verifier function that checks whether the
property is preserved in the regenerated program or \say{unknown}.
Each part of the plugin can produce logical atoms or rules which are fed to the \sys 
logical engine and can influence the regeneration process.

\section{Evaluation}

We evaluate \sys along three axes:

\begin{itemize}
  \item[\textbf{Q1}] \textbf{Correctness and Regeneration Quality}: Does \sys preserve program behavior and produce high-quality programs?
  \item[\textbf{Q2}] \textbf{Transformation Success}: Can \sys handle a diverse set of regeneration goals?
  \item[\textbf{Q3}] \textbf{Planning Efficiency \& Precision}: Does minimizing extracted properties help regeneration succeed faster and more reliably?
\end{itemize}

\subsection{Benchmarks \& Methodology}

We evaluate \sys on six benchmark suites, summarized in \cref{tab:benchmarks}.
These benchmarks include popular codebases from real-world package ecosystems, obfuscated or malicious components, and handcrafted microbenchmarks that reflect the use cases discussed in \cref{sec:example}.

\begin{table}[h]
\centering
  \caption{\textbf{Benchmark summary}. 
  Benchmark programs have been sourced from several sources.
  The $N$ column shows the number of programs/functions from each benchmar.
  }
\begin{tabular}{llrl}
\toprule
Benchmark                                               & Description                    & $N$ & Source \\
\midrule
  Rosetta Code                                          & Programming tasks              & 5 & \cite{rosettacode} \\
  npm utilities                                           & Utilities from npm             & 10 & \cite{regbench2025} \\
Python utilities                                        & Utilities from PyPi            & 1 & \cite{regbench2025} \\
  SSCA                                                    & Sneaky SSAs    & 3 & \cite{ev:eurosec:2022, es1, ohm2020backstabber} \\
  IOCCC                                                   & Obfuscated C code              & 2 & \cite{ioccc} \\
Microbenchmarks                                         &          & 3    & This paper \\
  \hspace{.5em} \ttt{flatmap-stream}                      & Secure regeneration            &      & \cite{es1}  \\
  \hspace{.5em} \ttt{Q_rsqrt}                             & C refactoring        &      & \cite{fast_inv_sqrt}  \\
  \hspace{.5em} \textsf{MusicDB}                          & Modularization           &      & \cite{codewithsadeemusicplayer} \\
\bottomrule
\end{tabular}
\label{tab:benchmarks}
\end{table}

Each task consists of an input program, a \sys query expressing desired properties, and a test suite or functional oracle.

\subsection{Q1: Correctness and Regeneration Quality}

We evaluate correctness using developer-made test suites 
shipped with the original software components as well as by manually inspecting
the regenerated source code and report on correctness and the other query target metrics.
When no test suite is available, we only do manual inspection.

\begin{table}[h]
  \centering
  \caption{Correctness: Percentage of tasks passing all I/O tests.}
  \begin{tabular}{lc}
    \toprule
    Benchmark & \sys Pass Rate \\ 
    \midrule
    Rosetta Code & \xxx/\xxx  \\
    npm Utilities & \xxx/\xxx \\
    Python Utilities & \xxx/\xxx \\
    SSCA & \xxx/\xxx  \\
    IOCCC & \xxx/\xxx  \\
    Microbenchmarks & \xxx/\xxx  \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Q2: Transformation Success}

We test whether \sys meets transformation goals across different categories.
Each row corresponds to a class of transformation query (e.g., removal of side effects, modular decomposition).
We count success when the generated program meets all stated properties and passes correctness checks.
\sys consistently succeeds at expressing non-trivial goals that GPT-4 baseline prompts often fail to satisfy.

\subsection{Q3: Planning Efficiency \& Precision}

To evaluate the impact of \sys's property planning, we compare it against a GPT-4 baseline that performs prompt-only rewriting without explicit access to guiding properties.

We consider two criteria for each regenerated program:

\begin{itemize}
  \item \textbf{Wanted Property Satisfaction:} Does the output program satisfy all positive properties declared in the query (\eg \texttt{pure}, \texttt{language(haskell)})?
  \item \textbf{Unwanted Property Avoidance:} Does the output avoid introducing disallowed behaviors (\eg side effects, global state, unsupported language features)?
\end{itemize}

\begin{table}[h]
  \centering
  \caption{
    Planning precision: Number of tasks where positive properties are satisfied ($P_p$) and negative properties are avoided ($P_n$).
  }
  \begin{tabular}{lcc}
    \toprule
    Variant & $P_{p}$ & $P_{n}$  \\
    \midrule
    \sys & \xxx & \xxx \\
    GPT-4 & \xxx & \xxx \\
    \bottomrule
  \end{tabular}
\end{table}

\sys consistently satisfies more of the declared properties while avoiding unintended behaviors. For instance, in the \texttt{flatmap-stream} benchmark, GPT-4 often preserved malicious side effects when the code appeared innocuous, whereas \sys generated provably pure rewrites by treating the source as untrusted and grounding regeneration in I/O behavior alone.

\heading{Planning overhead}
We measure the time taken by \sys's logic engine to select the minimal set of
properties required to satisfy each query. 
\begin{table}[h]
  \centering
  \caption{Planning overhead: time taken by \sys's logic engine to compute minimal property set.}
  \begin{tabular}{lcc}
    \toprule
    Benchmark & Properties Considered & $T_{\text{plan}}$ (ms) \\
    \midrule
    Rosetta Code & \xxx & \xxx \\
    npm Utilities & \xxx & \xxx \\
    Python Utilities & \xxx & \xxx \\
    SSCA & \xxx & \xxx \\
    Microbenchmarks & \xxx & \xxx \\
    \bottomrule
  \end{tabular}
\end{table}

Across all benchmarks, the planning step completes in under \xxx ms on average.
% TODO: re-add this
% adding negligible latency compared to synthesis time. Moreover, by minimizing
% the number of properties to extract and verify, planning reduces the total time
% and resource usage of the regeneration pipeline.

These results support the claim that \sys's planning logic---driven by explicit
property extraction and knowledge-base reasoning---yields safer and more
controllable transformations than prompt-only LLM-based systems.

\section{Related Work}

\heading{Program Synthesis}
Program synthesis research spans a rich space of techniques.
Classical techniques emphasize correctness and
interpretability, synthesizing programs from specifications~\cite{alur2013syntax, feser2015synthesizing, gulwani2011automating},
type constraints \cite{polikarpova2016program},
or examples \cite{jha2010oracle, raza2018disjunctive, singh2016blinkfill}.
These works often target narrowly
defined domains such as string manipulation~\cite{harp:ccs:2021} or data migration
\cite{yaghmazadeh2018automated} and focus on ensuring provable guarantees.
More recently, LLM-based methods~\cite{austin2021program, chen2021evaluating}
explore broad-domain code generation via prompt-based conditioning.
While remarkably flexible, these models are prone to hallucination and struggle
with reasoning under negation or satisfying structured constraints~\cite{xu2023llmfoolitselfpromptbased, wu2023deceptpromptexploitingllmdrivencode,jiang2024llmsdreamelephantswhen,hwang2024thinkpinkelephant}.

\sys leverages insights from both these paradigms.
Unlike purely symbolic or purely neural approaches, it leverages LLMs within a
constrained planning framework to guide transformations while remaining
responsive to property-based requirements. This hybrid design aims to combine
the expressiveness of LLMs with the controllability and robustness of symbolic
reasoning. % TODO: focus on that \sys focuses on this specific problem domain

\heading{Program Transformations and Refactoring}
The evolution and maintenance of software systems often involve transformations
such as refactoring, translation, or security hardening. Foundational work on
refactoring~\cite{Fowler99} and its taxonomy~\cite{Mens04} characterizes the
space of behavior-preserving modifications. More recent perspectives emphasize
developer experience and usability, particularly in the context of APIs~\cite{Myers16}.

Tools in this space have largely operated via syntax-tree transformations or
domain-specific templates. Verification-based methods such as \textsc{Dafny}
\cite{Leino10} and \textsc{CBMC} \cite{Clarke04} ensure functional correctness
but require heavy formalization.
Others automate migration via
programming-by-example or interactive guidance \cite{gulwani2017program, le2017interactive}.

Kvasir aims to generalize this landscape by treating transformations as
goal-directed regenerations—rather than prescriptive rewrites, driven by
properties such as program behaviors and architectural goals, expressed declaratively.

\heading{Logic Programming and Constraint Solving}
Logic programming has long served as a foundation for reasoning in programming
tools, offering a declarative model for expressing properties and constraints.
Answer Set Programming (ASP), in particular, provides a rich framework for
expressing defaults, exceptions, and optimization via stable model semantics
\cite{Gelfond_2000, Gelfond_2002, Eiter_2009}. ASP has been used for
declarative program analysis \cite{benton2007interactive}, automated planning
\cite{nguyen2020explainable, son2022answersetplanningsurvey}, and knowledge
representation in verification and synthesis pipelines.

\sys uses ASP as its planning back-end for the planning, decision-making
process, while also uses its straighftorward optimization capabilities.
For each transformation task, \sys formulates the property-satisfaction problem as a logic query and uses an ASP solver to
synthesize a feasible transformation plans.
This enables global reasoning across program components and enforces
user-specified constraints that are difficult for neural models to respect.
In addition, it offers a level of explainability that is useful in case transformations end up incorrect.

\bibliographystyle{ACM-Reference-Format}
\bibliography{bib.bib}

\end{document}