### 2025-06-20 docs/thesis/frontmatter/abstract.tex
Add thesis
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

\section*{Abstract}
\addcontentsline{toc}{section}{Abstract}

Modern software systems are constantly evolving—requiring developers to
  refactor legacy code, adapt to changing APIs, translate across languages, and
  patch security vulnerabilities. These transformations are often brittle,
  manual, and error-prone. Large language models (LLMs) offer new opportunities
  for automating such tasks, but their outputs are shaped by surface cues, lack
  semantic guarantees, and require brittle prompting strategies. This paper
  presents \sys, a system for controllable program regeneration that combines
  logic-based planning with LLM-guided synthesis.
  By expressing transformation
  goals as declarative constraints over program properties---such as input/output
  behavior, side-effect freedom, language, or modular structure, \sys
  regenerates code that satisfies explicit semantic requirements.
  We demonstrate \sys across a wide range of tasks, including security
  hardening, language translation, and idiomatic rewriting, and highlight its
  ability to deliver interpretable, high-quality transformations with minimal
  developer guidance.