### 2025-06-20 docs/thesis/frontmatter/frontpage.tex

\title{\thesistitle} 
\author{\thesisauthor} 
\date{\thedate}

\begin{titlepage}

\newgeometry{left=11mm,right=11mm,top=50mm,bottom=0pt}
\pagecolor{frontbackcolor}
\color{\frontpagetextcolour}

{ % Thesis title (to change see Setup/Settings.tex) 
\Huge
\begin{tabular}{p{\linewidth}}
\TitleFont{\thesistitle}   \\ 
\titlefont{\thesissubtitle} \bigskip \\ 
{\huge \documenttype} \bigskip \\
{\Large \thesisauthor~(\studentnumber)} \bigskip \\
\end{tabular}
}

% DTU department (to change see Setup/Settings.tex) 
\begin{tikzpicture}[remember picture,overlay]
\node[anchor=north east, 
      xshift=-10mm, 
      yshift=-12mm] 
      at (current page.north east) 
      {
        \color{\frontpagetextcolour}
        \begin{tabular}{r} 
        \textbf{\department} \\ 
        \departmentdescriber
        \end{tabular}
      }; 
\end{tikzpicture}

% DTU logo
\begin{tikzpicture}[remember picture,overlay]
\node[anchor=north west, 
      xshift=8.9mm, 
      yshift=-8.3mm] 
     at (current page.north west) 
     {\includegraphics[width=14.75mm,keepaspectratio]{figs/Logos/\dtulogocolour_\targetcolourmodel.pdf}}; 
\end{tikzpicture}

% Cover photo
\begin{tikzpicture}[remember picture,overlay]
\node[anchor=south, % anchor is bottom of picture
      xshift=0pt, 
      yshift=-2.9mm] % shifting picture to actually be at the bottom of the page
     at (current page.south) % placement at bottom of the page
     {\includegraphics[height=18.9cm,keepaspectratio]{figs/DTU_stock_photo.jpg}};
\end{tikzpicture}

\end{titlepage}